---
title: "Project: Exploring NYC Schools Survey Data "
author:"Ashish Kothari"
date:"August 13, 2020"
output: html_notebook
---

In this project, I will analyze data from the New York City school department to understand whether, parent, teacher and student perception of the following factors affect average school SAT scores:

1. Safety
2. Engagement
3. Communications
4. Academics

I'll start by installing the packages needed for analysis


```{r}
library(dplyr)
library(stringr)
library(purrr)
library(tidyr)
library(ggplot2)
library(readr)
```
Import the data into R
```{r}
combined<- read_csv("combined.csv")
survey<- read_tsv("masterfile11_gened_final.txt")
surveyd75<- read_tsv("masterfile11_d75_final.txt")
View(combined)
View(survey)
```
Filter the data

```{r}
survey_clean<- survey%>%
filter(schooltype == "High School")%>%
select(dbn:aca_tot_11)
```

```{r}
surveyd75_clean<- surveyd75%>%
  select(dbn:aca_tot_11)
```

Combining both survey_clean and surveyd75_clean
```{r}
survey_total<- survey_clean%>%
bind_rows(surveyd75_clean)
```

rename dbn to match with combine column name(DBN)
```{r}
survey_total<- survey_total%>%
  rename(DBN = dbn)
```

Joining survey dataset with combined(School data)

```{r}
combined_total<- combined%>%
  left_join(survey_total, by="DBN")
```

Create correlation matrix to see interesting relationships between pairs of variables
```{r}
cor_mat<- combined_total%>%
  select(avg_sat_score, saf_p_11:aca_tot_11)%>%
  cor(use= "pairwise.complete.obs")

cor_tib<-(cor_mat)%>%
  as_tibble(rownames="variable")

```
look for correlations of other variables with avg_sat_score that are greater than 0.25 or less than -0.25(stronger relationship)
```{r}
strong_cor<- cor_tib%>%
  select(variable, avg_sat_score)%>%
  filter(avg_sat_score>0.25 | avg_sat_score < -0.25)
```

Creating scatter plots to see the relationship in detail
```{r}
plots<- function(x,y){
  ggplot(data= combined_total)+
    aes_string(x=x,y=y)+
    geom_point(alpha=0.3)+
    theme(panel.background = element_rect(fill="white"))
}

x_var<- strong_cor$variable[2:5]
y_var<- "avg_sat_score"

map2(x_var,y_var,plots)
```
Reshaping the dataframe
```{r}
combined_survey_gather <- combined_total%>%
  pivot_longer(cols = saf_p_11:aca_tot_11,
               names_to = "survey_question",
               values_to = "score")
```

Create new variables,`response type` and `question` from the `survey_question` variable

```{r}
combined_survey_gather<- combined_survey_gather%>%
  mutate(response_type= str_sub(survey_question,4,6))%>%
  mutate(question= str_sub(survey_question,1,3))
```

Replace `response type` variable values with names "parent", "teacher","student","total" using `if_else()` function

```{r}
combined_survey_gather<- combined_survey_gather%>%
  mutate(response_type= ifelse(response_type=="_p_","parent",
                                ifelse(response_type=="_t_","teacher",
                                       ifelse(response_type=="_s_","student",
                                              ifelse(response_type=="_to_","total","NA")))))
```
Make a boxplot to see if there appear to be difference in how the three groups of responders(parents, students, and teachers) answered the four questions.

```{r}
combined_survey_gather%>%
  filter(response_type!= "total")%>%
  ggplot()+
  aes(x=question,y= score, fill= response_type)+
  geom_boxplot()
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

